### 2013-05-23 milestone.tex
Implemented cross validation for KNN. Began milestone writeup.

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\title{Exploration of Digit Recognition Algorithms}

\author{
Harnoor Singh \\
Department of Computer Science \& Engineering\\
University of Washington \\
\texttt{hsingh@cs.washington.edu} \\
\And
Brian Walker \\
Department of Computer Science \& Engineering \\
University of Washington \\
\texttt{bdwalker@cs.washington.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\section{Objective}

As described in our project proposal, we are using the Kaggle digit dataset to classify written numbers.

Our goal is to implement digit recognition using K Nearest Neighbors (KNN) and a Support Vector Machine (SVM) and compare the two approaches. A stretch objective is to enable data that isn't a part of the Kaggle dataset to work with our algorithms. This would require some form of normalization in the input image for consistency.

\section{Progress}

\subsection{KNN}

We have made significant progress in both the KNN algorithm and the SVM algorithm. Our KNN algorithm is completely implemented as is cross validation for it.

To compute the K nearest neighbors, we take the euclidean difference between the feature we are classifying, $X_1$, and every feature in our training set.

\begin{center}
$\sum\limits_{X_2 \exists N} {(X1 - X2)^2}$
\end{center}

Once we have calculated the K closest features, a simple voting method is used to classify $X_1$.

In order to select a value of K we do cross validation on a subset of the training set. We begin by passing a list of potential K values we are considering. For each item, $K_i$ in this list, we hold out $K_i$ samples from our training set and construct our classifier. We then attempt to classify the $K_i$ samples and record the error rate.

The K value that led to the lowest error rate is used to build our final classifier.

\subsection{SVM}

TODO

\begingroup
\renewcommand{\section}[2]{}%
%\renewcommand{\chapter}[2]{}% for other classes
\bibliographystyle{unsrt}
\bibliography{references}	
\endgroup

\end{document}